---
title: "SDM2_HW2"
author: "Sri Pavankrishna Yenugu"
date: "2024-03-08"
output:
  html_document: default
  pdf_document: default
---

2)
Consider the California Housing Data from KAGGLE. You are going to apply
association rules to this data. (https://www.kaggle.com/camnugent/california-
housing-prices).
a. Read in the data and transfer to a binary incidence matrix. Visualize this matrix.

```{r}
setwd("/Users/sripavanyenugu/Downloads")

library(arules)

# Read the data
my_housing_data <- read.csv("housing.csv")

my_housing_data$ocean_proximity <- as.factor(my_housing_data$ocean_proximity)

my_numeric_cols <- c("housing_median_age", "total_rooms", "total_bedrooms", 
                  "population", "households", "median_income", 
                  "latitude", "longitude")

my_housing_data[, my_numeric_cols] <- lapply(my_housing_data[, my_numeric_cols], as.factor)
my_housing_data$median_house_value <- cut(my_housing_data$median_house_value, breaks = 5)

# Convert to binary matrix
my_housing_trans <- as(my_housing_data, "transactions")
my_housing_rules <- apriori(my_housing_trans, parameter = list(support = 0.001, confidence = 0.5))

inspect(head(my_housing_rules))


```
b. What are the top three high lift rules?

```{r}
my_trans <- as(my_housing_data, "transactions")
my_rules <- apriori(my_trans, parameter = list(support = 0.01, confidence = 0.5))
top_three_lift_rules <- inspect(sort(my_rules, by = "lift", decreasing = TRUE)[1:3])

```

c. What are the top 4 rules according to confidence?

```{r}

inspect(sort(my_rules,by="confidence",decreasing=TRUE)[1:4])

```
d. A person comes to you and wants to purchase an average priced home as close to the ocean as possible. What can you recommend and/or what can the expect? Use association rules to guide you.

Recommendations:
-For an average-priced home between $14,500 to $112,000, expect properties INLAND.
-For homes priced between $112,000 to $209,000 and with median ages 35 or 36, expect properties near the ocean.
These recommendations are based on association rules from the dataset.


e. What characteristics in the data associate with low population areas?

INLAND Locations:
-Lower median incomes $0.5 to $1.75 are associated with INLAND locations.
-Lower median house values $14,500 to $112,000 are also associated with INLAND areas.
-Younger median housing ages (around 28) are linked to INLAND areas.

Low Population Areas:
-Lower median incomes, lower house values, and younger housing ages are characteristics associated with low population areas in the dataset.
-These associations suggest that INLAND locations, with their lower median incomes and house values, tend to have lower population densities.

